### 2025-03-07 setup.qmd

---
title: "Setup"
filters:
  - highlight-text
format: 
  html: default
  pdf: default
---

This works takes NM CTFM profiles, resamples them and calculates number of
flight present per hour, `H`, of date, `D` in H3 hexagonal cells
(at resolution `R`) covering the EUROCONTROL Member's airspace.

R and Python scripts are in the `R/` and `python/` directories respectively.

We use the following values:

* H3 hexagons at resolution `R = 2`: 
  - Average Area (km2): `86801.78`
  - Average edge length (Km): `182.51`
* H3 hexagons at resolution `R = 3`: 
  - Average Area (km2): `12393.43`
  - Average edge length (Km): `68.98`
* bounding box approximating EUROCONTROL Area (decimal degrees):
  - xmin = -25.488281
  - ymin = 26.638253
  - xmax = 45.407181
  - ymax = 71.864780
* bounding box containing the HEX at resolution 2 that match
  the bounding box approximating EUROCONTROL Area:
  - xmin = -27.78166
  - xmax =  48.56037
  - ymin =  24.90098
  - ymax =  72.85065

::: {#fig-cells}

```{r cells-map}
#| out-width: "100%"
#| echo: false
knitr::include_graphics("figures/cells_2.png")
```

H3 **[cells]{colour="#ff0000"}** ([red]{colour="#ffffff" bg-colour="#ff0000"})
at resolution 2 and
their **[centers]{colour="#0000ff"}** ([blue]{colour="#ffffff" bg-colour="#0000ff"}) covering EUROCONTROL Area.

:::

## Data Preparation

### Data Extraction
We first extract data for data `D` (for example `2024-08-01`)
1. flight list for date `D`
1. point profile for flown trajectories, i.e. M3/CTFM for date `D`
   in the correct bounding box.
   We use `point_profiles_tidy()` from the `{eurocontrol}` package

This is achieved via script `export_data.R`.
The relevant trajectories are saved in
`data-raw/trjs_{D}.parquet`, i.e. `data-raw/trjs_2024-08-01.parquet`.

### Trajectory Resampling
We load the trajectories and resample them at 30s using the Python `traffic` library.
This is performed by the `resample_aug.py` and `resample_dec.py`
scripts and is time consuming (40 min each).
Execute as:

> $ python resample_aug.py

The result is in (52min 51s)
`data/trajectories_2024-08-01_resampled_30s.parquet`

### Fast H3 Labelling
The script `prepare_pre_density.R` uses DuckDB for fast labelling
of each lon/lat to the relevant H3 cell (at resolution `R`)
and filtering for samples between bbox.
Keep only point in the hex cells of the bbox and for day `D`.
The result is saved in `trajectories_<YYYY-MM-DD>_resampled_30s_bbox_res_3.parquet` for date `D`.

## Cell Occupancy and Density

The script `prepare_density.R` calculates the occupancy per hour `H` of `D` 
for each cell at resolution `R`
.