### 2025-06-09 paper.qmd
spell-checking

---
title: "Preparing for Potential Closure of European Airspaces due to Re-entering Space Objects"

keywords: 
  - space debris
  - airspace closure
  - aviation safety

filters:
  - acronyms
  - highlight-text
  - abstract-section

crossref:
  eq-prefix: ""
  fig-prefix: Fig.   # (default is "Figure")
  fig-title: Fig.
  # tbl-prefix: table    # (default is "Table")
  ref-hyperlink: false # (default is true)

format:
  ieee-pdf:
    keep-tex: true
    colorlinks: false
    # fig-pos: 'H'
    journaltype: conference
    papersize: a4
    include-in-header:
      - file: media/preamble.tex
    include-before-body:
      - text: |
          \let\qty\SI
          % NOTE: make siunitx's \qty work without clashing with 
          %      the one in physics
  ieee-html: 
    html-math-method: mathjax
    include-in-header:
      file: media/siunitx.html
author:
  - id: es
    name: Enrico Spinielli
    affiliations:
      - ref: aiu
    orcid: 0000-0001-8584-9131
    email: enrico.spinielli@eurocontrol.int
    attributes:
      corresponding: true
    role: 
      - conceptualization
      - writing
      - software
  - name: Rainer Koelle
    affiliations:
      - ref: aiu
    orcid: 0000-0003-2810-1404
    email: rainer.koelle@eurocontrol.int
    role: 
      - editing
  - name: Quinten Goens
    affiliations:
      - ref: aiu
    orcid: 0000-0003-2187-1924
    email: quinten.goens@eurocontrol.int
    role: 
      - editing
affiliations:
  - id: aiu
    name: EUROCONTROL
    department: Aviation Intelligence Unit
    city: Brussels
    country: Belgium
  - id: fake
    name: ACME
    department: Best Tricks
    city: Los Angeles
    country: USA
  - id: empty

bibliography: media/references.bib
acronyms:
  insert_loa: false
  fromfile:
    - ./media/acronyms.yml
editor: 
  markdown: 
    wrap: 72
---
```{r}
```

```{r}
```

# Abstract

<!-- teaser -->

The recent decades observed an increase in the number of space activities.
This resulted in higher numbers of re-entries at the end of the respective 
mission lifespan or orbital decay. The risk of collision between uncontrolled
re-entering space debris and aircraft is increasing due to the continual 
growth of flights and of orbiting bodies and their relevant re-entries. 
<!-- why important --> National and international
authorities need to balance safety and economic impact of adopted
measures. This is particularly challenging should a re-entry event be announced on short notice.
Thus, there is a need for effective tools to reduce the uncertainty and increase the safety awareness of uncontrolled re-entries. 
<!-- approach / experiments --> To determine the risk exposure, this paper follows a data-driven approach.
We employ hourly and daily spatial flight densities calculated from trajectories
as recorded by EUROCONTROL's Network Manager and binned with the Uber H3
geospatial indexing system.
Impact risk probabilities are then calculated based on previous theoretical
work. 
<!-- results / conclusions --> This work enhances the previous model by using
actual flown flight trajectories as a basis for time-varying flight densities
rather than relying on proxies such as the population density or ADS-B/Mode S
message counts. These densities combined with a buffered area along the eventual path of re-entry, can provide mid-air impact probabilities to authorities
as an aid for deciding on safety measures such as airspace closures and/or
re-routings.
The findings and the approach are demonstrated on the basis of an expected re-entry from an earlier space mission.\
The approach in this paper enables a computing time efficient mechanism to support the re-entry related decision-making.

# Introduction

<!-- teaser opening --><!-- might need a rewrite - first jotting down -->
The quest for space is a driver for technological development and exploitation.
Key milestones include the first human travel to space and circumnavigating the globe, the first and now many succeeding successful landings on the moon, and regular operations of space stations in the orbit.
These developments helped to identify options for the wider use of space, associated technological or operational benefits, and its commercialisation.
This is showcased by an increasing number of activities aimed to generate value through the exploration of space ranging from scientific understanding and experimentation, sustainable production of goods in a non-gravity environment, space-asset based services, and recently even leisure.
The increased number of launches, however, also increased the number of artefacts <!-- falling down and --> re-entering the atmosphere.  
While ascent operations are typically well controlled and coordinated, a high number of re-entries ensue based on lifespan, decay, or side-effects, such as debris, failed components, etc.

<!-- zooming in on problem -->

With forecasts predicting a steadily growing number of flights, see
@fig-forecast, and the modern race to commercialise space, there is a
need to re-evaluate the risks of collision and casualties due to
uncontrolled debris re-entries. The increasing risk landscape poses
challenges in terms of oversight and risk management. The operational and 
economic demand by air transport requires a trade-off with large-scale
airspace and flow management measures to address the inherent safety risk.
There is a growing demand by overseeing authorities for methodologies and
tools to assess which safety measures are the most adequate to minimise
potential casualties and economic impact.

The increase in space operations does not exist in a vacuum. The
continual growth of air traffic - despite temporal bumps - is evidenced
by the surging number of flights in Europe [@forecastUpdate2025] and
across the world. On-going development/deployment activities of
mega-constellations of satellites in \acr{LEO}, such as the
already active Starlink and OneWeb or planned ones (e.g. Kuiper,
Starshield, Xingwang), will add thousand of orbiting objects,
c.f. [@tbl-megaconstellations;@megaconstellations] with a short
operational life. Some of these will inevitably re-enter the Earth's
atmosphere in an uncontrolled manner.
These numbers and occurrence rates will pose challenges to the steadily increasing air traffic.

<!-- now we have the risk / topic established -- next we need something to speak about a "high-level goal" or "what other work looked into to place this paper as a 'further development' -->

Aviation safety risk management is a well-established domain in air
transportation. Despite the increasing need for situational awareness of
space objects and their re-entry, the classical risk avoidance paradigm
revolves around restricting the airspace. 
Such airspace closures impact the flow of air traffic, as the operating schedule
will get distorted due to delays or re-routings circumnavigating the exposed volume of airspace.
While the probability of a strike is low, it can result in catastrophic 
consequences. 
The growing number of space operations requires therefore tools to help improve
the monitoring and impact of uncontrolled re-entries, techniques to help reduce
the uncertainty in terms of the dimension and duration of required airspace
measures, including computationally efficient and fast methods to determine the risk exposure.

<!-- contribution of the paper -->

This paper develops a practical and straightforward data-driven pipeline
to rapidly assess the risk from uncontrolled re-entry of space debris
for the European region. Apart from the general risk assessment of
potential re-entries, a soon to occur re-entry of debris from a 1972
failed USSR mission to Venus is analysed.

::: {#fig-forecast}
```{r forecast}
#| out-width: "100%"
#| echo: false
#| fig-scap: "Forecast"
knitr::include_graphics("media/figures/forecast_statfor.png")
```

EUROCONTROL 7-year forecast 2025-2031 (Spring 2025).
:::

```{r}
#| label: tbl-megaconstellations
#| tbl-cap: Mega-constellations (planned > 1000) as per filing to the \acr{ITU}
#| tbl-cap-location: bottom

library(tibble)
library(gt)

megas <- tibble::tribble(
  ~constellation, ~state, ~launched, ~active, ~planned, ~first_launch,
  "Starlink", "US", 4714, 3521, 4714, 2018,
  "Starlink2A", "US", 3689, 3070, 6720, 2022,
  "Starlink2", "US", 0, 0, 30456, NA,
  "OneWeb", "UK", 660, 635, 716, 2019,
  "OneWeb2", "UK", 0, 0, 2304, NA,
  "Kuiper", "US", 29, 27, 3232, 2023,
  "StarShield", "US", 193, 126, 32, 2022,
  "Xingwang", "CN", 50, 10, 996, 2021,
  "Qianfan", "CN", 90, 28, 32, 2024,
  "Guangwang", "CN", 0, 0, 12992, NA,
  "Yinhe", "CN", 8, 7, 1000, 2020,
  "Hanwha", "KR", 0, 0, 2000, NA,
  "Lynk", "US", 10, 6, 2000, 2020,
  "Astra", "US", 0, 0, 13620, NA,
  "Telesat", "CA", 0, 0, 300, NA,
  "HVNET", "US", 0, 0, 1440, NA,
  "SpinLaunch", "US", 0, 0, 1190, NA,
  "Globalstar3", "DE", 0, 0, 3080, NA,
  "Honghu-3", "CN", 0, 0, 10000, NA,
  "Semaphore", "FR", 0, 0, 116640, NA,
  "E-Space", "US", 4, 0, 337323, 2022
)

# ~constellation, ~state, ~launched, ~active, ~planned, ~first_launch,
megas |> 
  gt() |> 
   cols_label(
    constellation ~ "Constellation",
    state ~ "Country",
    launched ~ "Launched",
    active ~ "Active",
    planned ~ "Planned",
    first_launch ~ "First launch"
  ) |> 
   sub_missing(
    columns = "first_launch",
    missing_text = "-"
  ) |> 
  opt_table_font(size = "10px")

```

# Method

<!-- ESTABLISH SOME CONTEXT ---------------------------------------------- -->

The modelling and tracking of uncontrolled re-entries poses challenges in terms
of safety awareness and operational risk.
Actual and historic data is sparse, and the size, shape, and attitude or
rotation of re-entering objects increase the level of uncertainty for
modelling. To support the decision-making, common approaches employ a mix of
handling the uncertainty of
object or debris movement and the associated exposed area for encounter,
including temporal uncertainty windows.
Thus, computationally efficient methods to assess and continuously revalidate
the collision risk of uncontrolled re-entries and aircraft will help to
fine-tune the decision-making and geospatial and temporal constraints on
aviation.

<!-- is there more to say and provide context? -->

<!-- APPROACH / METHOD --------------------------------------------------- -->

<!-- it would be cool, if we can add a diagram here that shows the path from data, major processing building blocks to the results -->

<!-- as a minimum, we should say something about the WHAT -->
In general, the method for estimating the risk to commercial aircraft
from re-entering debris objects revolves around estimating the exposure
of a single flight and the population of flights to the re-entering
object. In that respect, a robust determination of the flight density for encounter airspace volumes is a
key aspect.

::: {#fig-cells}
```{r cells-map}
#| out-width: "110%"
#| echo: false
#| fig-scap: "H3 cells at resolution 3"
knitr::include_graphics("media/figures/cells_3.png")
```

H3 cells (red) at resolution 3 and their centres (blue) covering the
bounding box of 43 EUROCONTROL Member States \acr{FIR}s (blue shaded
polygon).
:::

We follow the same mathematical approach of [@patera2008;
@wrightAirspaceClosuresDue2025]. However, this paper augments the
approximation of risk exposure. Instead of approximating flight density
with population density [@patera2008] or aircraft transponder data[^adsb]
[@wrightAirspaceClosuresDue2025], we use actual flown flight
trajectories.
This provides a more realistic figure of the number of affected air traffic.
Thus, flight density is directly related to the number of flight operations per 
volume of airspace.

[^adsb]: Crowdsourced ADS-B/Mode S data suffer from lack of coverage away from landmasses and from the need to filter away position reports for stationary aircraft at airports which unnecessarily inflate the density at the location.

This work focusses on the European region. The hourly flight density
estimation is based on _flown_ \acr{IFR} trajectories as recorded
by EUROCONTROL's \acr{NM}[^1]. These trajectories are reconstructed from
the filed \acr{FP} and recalculated when big deviations are detected from
position reports received from \acr{ATC}.

To decrease spatial complexity, the data is
binned using Uber's H3 geospatial indexing system [@uber-h3]. The
reduction of the spatial complexity enables fast processing. For this
work, we employ a H3-cell-resolution of 3. This corresponds to hexagons
of an average area $\qty{12393}{km^2}$ and an average edge length of
$\qty{68.979}{km}$ [@h3resolutions] as depicted in @fig-cells.

The choice of hex resolution is a compromise between having to use
highly detailed trajectories (more 4D positions and more processing and
storage) and capturing long enough portions of flight inside the cell.

To build an upper threshold for the flight impact exposure, this study uses the
busiest week during summer 2023 and 2024.
This ensures a reasonable value, as peak airspace utilisation provides
a handle on the air traffic operations and interactions of the different
aircraft as part of wider *'traffic flows'*.
The flight densities are characterised on an hourly basis for all days
in these weeks per H3-cell.

[^1]: NM coordinates \acr{ATFCM} operations on behalf of EUROCONTROL's 42 Member States and 2 Comprehensive Member States covering almost all \acr{ECAC} States, see blue polygon in @fig-cells.

::: {#fig-density-summer2024-res-3}
```{r density-winter-res-3}
#| out-width: "105%"
#| width: "9cm"
#| echo: false
knitr::include_graphics("media/figures/flight_density_2024-07-05_3.png")
```

Flight density in the EUROCONTROL Area on the 5th of July 2024 (the busiest day of the busiest summer week in 2024) at H3 cells resolution $3$. The black hexagon in the South of UK marks the densest cell at 24.56 flights/hour, just around the airport of London Heathrow.
:::

@fig-density-summer2024-res-3 shows an example of binning the hourly flight
density on the 5th of July 2024 using H3 hexagons.
The chart depicts the expected outcomes in terms of major
traffic flows, and areas of high or low traffic. For example, cells with
a density of zero are clearly visible for areas of war or no-fly zones
(Ukraine, Libya or Syria). At the same time, the major traffic flows (North Atlantic,
Canaries and East-bound over Türkiye) are clearly highlighted showing
densities of more than 10.

The next sections will develop the mathematical background for the
evaluation of the collision risk for which air traffic density is
a key component.

## Impact probability for a single flight and single debris

Safety and collision risk awareness is directly linked to the probability of a re-entering space object colliding with a traversing flight.
This probability, $p_k$, of a debris (point) object $k$, impacting an
aircraft (point) object over the area, $a$, is expressed as:

$$
p_k = \varepsilon_k \int_a \rho(\vec{x}) \sigma_k(\vec{x}) da
$$ {#eq-impact-probability-general}

where

* $\rho({\vec{x}})$ is the probability density of an aircraft in $\vec{x}$,
* $\sigma(\vec{x})$ is the impact probability density of a debris object $k$
in the position $\vec{x}$, and 
* $\varepsilon_k$ is the effective exposed area, i.e., the effective risk area for the $k^{th}$ debris object impacting
the aircraft (see @sec-eea later on).

Our approach builds on the discretisation of the geospatial risk area.
For this purpose we use a hexagon-based tessellation pattern.
Uber's H3 method ensures non-overlapping hexagons (and 12 pentagons), with an equidistant relationship between all neighbours, and enables an efficient indexing of the cells. 
While this paper focusses on the determination of the risk exposure per cell, the approach would allow to identify adjacent cells to account for the inherent uncertainty and risk appetite of decision-makers.

Using H3 to discretise the area of interest $a$ gives

$$
p_k = \varepsilon_k \sum_h \rho(h) \sigma_k(h).
$$ {#eq-impact-probability-h3}

The impact probability $p_{k}(h)$ for each cell $h$ is determined by the following expression

$$
p_{k}(h) = \varepsilon_k \rho(h) \sigma_k(h).
$$ {#eq-impact-k-debris-cell}

## Aircraft effective exposed area {#sec-eea}

For the actual collision encounter between a re-entering object and an aircraft, we assume the aircraft impact area to be much larger than the debris object area.
Thus, the area spanned by the debris object is smaller than the aircraft dimensions. 
This assumption holds for practically all commercial aircraft within the scope of this study, and we can neglect the debris impact area.
Accordingly, we can also simplify the geometry of the aircraft and model it as a rectangular solid with dimensions of wingspan ($W$), length ($L$) and height ($H$).

Regarding the aircraft-debris encounter geometry, this yields to two conceptual considerations:

1. if the aircraft were stationary, the area exposed to the debris impact would be the top face of the rectangular solid, and
2. if the debris were stationary, it would be the front face.

Our focus is on commercial aircraft in flight, and --- implicitly --- operating at higher altitudes (i.e., flight levels).
The en-route portion of a flight is characterised by the aircraft in straight and level flight, moving (mainly) horizontally.
At the same time, the trajectory of a falling re-entering debris object at regular en-route flight levels will be a vertical descent movement[^ballistic_descent].
This encounter geometry allows to express the effective exposed area, $\varepsilon_k$, as

$$
\varepsilon_k = \frac{s_{a} a_{F} + s_{d_k} a_{T}}{s_{d_k}}
$$ {#eq-effective-exposed-area}

where $s_{a}$ is the average aircraft speed, 
$s_{d_k}$ is the average fall speed of the $k$-th debris object, 
$a_T$ area of the aircraft as viewed from the top, and 
$a_F$ area of the aircraft as viewed from the from the front of the aircraft.

This can be further instantiated for specific aircraft types. In the
calculations the following is assumed to be true ($t$ subscript for
aircraft type)

* constant debris fall speed $s_{d_k} = \qty{145}{mph} = S$ 
  ([we adopt this value from pages 7-8 in @patera2008])
* constant aircraft speed, $s_{a} = C_t$, equal to its cruise speed, under the
  simplification that the majority of the flight time is spent cruising, 
* aircraft area as viewed from the top, $a_T = W_t \times L_t = T_t$, 
  where $W_t$ is the wingspan, and $L_t$ the length for the specific aircraft
  type $t$
* aircraft area as viewed from the front, $a_F = W_t \times H_t = F_t$, 
  where $W_t$ is the wingspan, and $H_t$ the height for the specific aircraft
  type $t$

The effective exposed area now only depends on the aircraft type $t$
(i.e., its dimensions and cruise speed) and can be rewritten as:

$$
\varepsilon_t = \frac{C_{t} F_{t} + S T_{t}}{S}
$$ {#eq-effective-exposed-area-simplified}

These aircraft properties are well-known design parameters. 
While there does not exist a single openly available date base for these aircraft parameters, we collected the information from a series of sources and integrated them into a lookup dataset.
The values $W_t$, $H_t$, $L_t$ and $C_{t}$ have been retrieved from [@performance_database;@skybrary;@doc8643]; we used $\varepsilon_t = \qty{500}{m^2}$
(and $pax = 7$)[^pax] for missing entries which usually are due to smaller aircraft not listed in the above sources.

[^pax]: `pax` values, i.e. number of passengers, will be used for future work on
        casualty risk calculation.

The impact probability for cell $h$ under consideration of a specific
aircraft of type $t$ becomes:

$$
p_{t,k}(h) = \varepsilon_t \rho_t(h) \sigma_k(h)
$$ {#eq-impact-k}

[^ballistic_descent]: *Even for a ballistic coefficient of \qty{50}{lbs/ft^2},
the flight path angle of a re-entering object at \qty{40000}{ft} is 
\ang{-86.6}, which corresponds to nearly vertical descent.
Since most debris objects have ballistic coefficients equal to or 
less than \qty{50}{lbs/ft^2}, the vertical descent assumption is a very
accurate approximation.* [@patera2008]

::: {#fig-weighting-inclination-i}
```{r weighting-inclination-i}
#| out-width: "90%"
#| echo: false
knitr::include_graphics("media/figures/weighting_inclination_i.png")
```

Impact probability density $\sigma(\varphi, i = 51.95)$ for debris of
inclination $i$ and altitude $\qty{550}{km}$.
:::

## Impact probability density for debris of inclination $i$

Debris that re-enter the Earth’s atmosphere via drag-induced decay will fly
circular orbits prior to re-entry even if initially on elliptical
ones.[^circular-orbit]
A characteristic point for the determination of the re-entry position is the *'ascending node'*.
This denotes the intersection point of the object crossing the plane of Earth's equator if moving from the southern hemisphere to the northern one.
The respective angular position, $\theta = \omega + \nu$, of the re-entry point from the
ascending node, see @sec-orbitals, is uniformly distributed for uncontrolled
re-entries, $\sigma(\theta) = 1 / (2\pi)$.
This is due to uncertainties in atmospheric drag and the gravitational effects of the
Equatorial bulge.
The height of the atmosphere is not a constant across the globe. Next to the gravitational effects, the atmosphere may also expand due to solar activity.
Given these uncertainties, we assume that the impact longitude, $\lambda$, is uniformly distributed.

The impact probability density, $\sigma(\varphi, i)$,
at latitude $\varphi$ of a re-entering object whose orbit has inclination $i$
(see @fig-orbital-elements) is given by ([Equation 15 in @patera2008]):

$$ \sigma(\varphi, i) = \frac{1}{2 \pi^2 R^2 \sqrt{(\sin^2(i) - \sin^2(\varphi))}}$$ {#eq-impact-probability-density}

where $R$ is the distance from the centre of the Earth of the object potentially
being impacted by the debris.

::: {#fig-weighting}
```{r weighting}
#| out-width: "100%"
#| echo: false
#| fig-pos: H
knitr::include_graphics("media/figures/weighting.png")
```

Weighting function derived from 25 years of re-entries as an empirical
approximation of $\sigma(\varphi)$.
:::

So for example if a building is located at latitude \ang{30} and has an area 
of $\qty{1000}{m^2}$ and a piece of debris re-enters the Earth’s atmosphere
from an orbit having an inclination of \ang{45}, the impact probability is 
given by

$$
\begin{aligned}
P &= \Delta A \times \sigma(\varphi, i) = 1000 \times \sigma\Bigl(\frac{\pi}{6}, \frac{\pi}{4}\Bigl) \\
  &= 1000 \times \num{2.490756e-15} = \num{2.490756e-12}
\end{aligned}
$$ {#eq-impact-probability-example}

where we used the equatorial value of $\qty{6378}{km}$, for the Earth's radius,
$R$, as the distance of the building from the centre of the planet.

The impact probability density $\sigma(\varphi, i)$ for an inclination $i$ of
$\ang{51.95}$ is plotted in @fig-weighting-inclination-i and takes a characteristic U shape with zero values for latitudes beyond $i$.
This density has been derived by simulating a full circular orbit as detailed
in @sec-orbitals.

[^circular-orbit]: In fact drag at perigee will lower apogee and similarly drag
at apogee will lower perigee. But because drag at perigee is higher due to
denser atmosphere (the perigee is closest to Earth), the apogee  will decrease faster than perigee and the orbit will then tend to become a circular one, see figure about apogee/perigee tracking in [@kosmos482reentry] for empirical evidence.

## Impact probability density for debris

Considering historical uncontrolled debris re-entry data, we can numerically
derive the debris impact probability as a weighting function by summing up and
averaging all single densities:

$$
\sigma(\varphi) = \frac{1}{N} \sum_{k = 1}^N \sigma(\varphi, i_k)
$$ {#eq-debris-impact-probability-historical}

From [@gcat] we extracted 1968 debris re-entries during 25 calendar years from 
1st Jan 2000 till 31st Dec 2024.
The result of combining all the U-shaped densities can be seen in @fig-weighting.

Spatially projecting $\sigma(\lambda,\varphi)$ on the H3 cells in the
\acr{NM} area, i.e. taking the mean of all the half degree weighting values
and spreading in longitude,
we obtain the weightings per cell (per hour $H$), $W_H(h)$, as in
@fig-weighting-nm-area.

![Hourly weightings for H3 cells, $W_H(h)$, in \acr{NM} area.](media/figures/weighting_on_Earth.png){#fig-weighting-nm-area out-width=100%}

Then (@eq-impact-k) becomes

$$
p_t(h) = \varepsilon_t \rho_t(h) \sigma(h).
$$ {#eq-impact-cell}

## Traffic density per cell

In order to derive $\rho_t(h)$, we have taken the trajectories of flown flights
in the \acr{NM} Area, resampled at $\qty{30}{s}$ so that the calculation of the
occupancy, i.e. the time spent in the hex, per aircraft type is just a matter
of counting the number of points per cell.
This binning operation is extremely fast to perform in Uber's H3 hierarchical
geospatial indexing system and a real advantage in processing the huge amount of
data point representing the flight trajectories.

For July, $5^{th}$ 2024 alone, we extracted $37721$ flights with
$33,482,686$ 4D positions (at $\qty{30}{s}$ sampling), where 19 aircraft types
cover each at least 1% of the total and the remaining 266 (!) types account
for the residual 20.8%, see @fig-aircraft-types.

## Hourly impact probability per H3 cell

When combining the hourly ($H$) aircraft occupancy (per aircraft type) with the
relevant weightings per $h$ cell, $W_H(h)$, and the effective exposed
area, $\varepsilon_t$, we can numerically compute the
hourly impact probability for each hexagonal cell as:

$$
 E_{h, H} =  \sum_t W_H(h) \cdot \rho_{t,H}(h) \cdot \varepsilon_t
$$ {#eq-impact-probability-hourly}

where $E_{h,H}$ is the collision expectation in the hexagon $h$ at hour $H$.

::: {#fig-aircraft-types}
```{r aircraft-types-july-2024}
#| out-width: "100%"
#| echo: false
knitr::include_graphics("media/figures/aircraft_types.png")
```

Aircraft types on 2024-07-05 in \acr{NM} Area. 
:::

@fig-collision depicts the determined hourly collision expectation. As
identified in @fig-density-summer2024-res-3 areas of higher traffic
density observe a higher collision expectation.
<!-- WHAT ELSE CAN WE SAY ABOUT THIS PLOT? -->

::: {#fig-collision}
```{r collision}
#| out-width: "100%"
#| echo: false
knitr::include_graphics(here::here("media", "figures", "collision_expectation_2024-07-05_3.png"))
```

Averaged hourly collision expectation on 2024-07-05.
:::

# Use Case: KOSMOS 482 Descent Craft re-entry

While this paper was being written, the KOSMOS 482 Descent Craft was being
monitored for an uncontrolled re-entry, forecast at 06:39 May 10, 2025
($\pm 1.5$ hours) [@kosmos482reentry].

![Venera-4 lander mock-up at the Memorial Museum of Cosmonautics. The Kosmos 482 Descent Craft is probably similar.](media/venera-capsule-in-museum.jpg){#fig-kosmos width=80%}

This is the lander module from a 1972 failed USSR Venera mission
[@SpaceReviewKosmos], which is reported to weigh about $\qty{480}{kg}$ and
to have a diameter of approximately $\qty{1}{m}$.
Being built to survive Venus's atmosphere which is much denser and hotter than
Earth's one (surface pressure 93 times that of Earth and temperatures reaching
\ang{467}C (\ang{872}F)), it could re-enter Earth intact and not break into
smaller fragments some of which could burn up.

Knowing that the debris inclination is $\ang{51.95}$, 
see @fig-weighting-inclination-i, we can derive the maximum
collision expectation using data for the busiest day in the busiest 2024 week
as a proxy for the traffic density on the day of re-entry.
Using the approach in this study and the weightings for this specific debris,
the cell with the highest impact risk is cell, `83195dfffffffff` 
(located North-West of London and overlapping with the airport of London Heathrow), with a daily value of probability collision risk of $2.63 \cdot  10^{-10}$.
This corresponds to an annual probability of a collision between this rocket
body and an aircraft of $9.6 \cdot  10^{-8}$ (1 in 10420000).
The risk is very low, for a comparison the odds of a person being struck by
lightning in a given year are less than one in a million
[@cdcLightningStrikeVictim2024].

Radar systems over Germany saw the descent craft at approximately 04:30 UTC and
06:04 UTC on May $10^{th}$. Given that the descent craft was not spotted at
the next expected pass at 07:32 UTC, it is likely that the re-entry already occurred before.

In general case when the re-entry track for a debris would be known
with some lead time, the H3 cells around, say,  \qty{35}{km} on each side
could be identified to retrieve the values of collision expectancy.
It would then be possible to decide whether it would advisable to close
the airspace and re-route traffic around the affected area.

::: {#fig-data-densities}
```{r densities}
#| out-width: "70%"
#| echo: false
#| fig-pos: t
knitr::include_graphics("media/figures/cubes.png")
```

Data cube of hourly flight densities data over H3 hexes (in Germany as
an example) for multiple dates, $d_1 \cdots d_n$.
:::

# Data and Code for Reproducibility

We have calculated collision risk probabilities for the two
busiest weeks in 2023 and 2024. The relevant dataset represents a data cube
as logically depicted in @fig-data-densities.

The processing of data for 1 day takes less than 30 minutes from
data extraction to impact probability calculation on a
high-end laptop (MacBook Pro M1 Max).

Code and sample data are available in the companion GitHub repository of
the paper at <https://github.com/euctrl-pru/flight_density_space_debris>.
We have also developed an R package `aviodebris` to heavy-lift the
various data processing steps which  is available at
<https://github.com/euctrl-pru/aviodebris>.

# Conclusion and Future work

<!-- opening statement linking back to abstract/intro and motivation for this work --------->
Space objects and debris re-entry is an increasing risk to aviation and the
general public.
The higher pace of commercial space operations will see an increase of launches 
and subsequent re-entries.
Dependent on the mission, objects re-entering the atmosphere can take on
different forms, shapes, and sizes.
On top, a higher number of these entries may be uncontrolled.
With the continual growth of air traffic, there is a need to enhance aviation
safety risk awareness and respond in a more tailored fashion to accommodate
the associated uncertainties, reduce the impact on aviation, and limit the risk
to the travelling public.

<!-- what was shown in the paper ------------------------------------------------------------>

This paper expands on previous work. With a focus on Europe, we establish an
upper bound for the traffic density based on actual flights.
In particular we derive impact risk probability for the European airspace
using flown trajectories.

The findings are applied to a use-case: the imminent re-entry of the
lander module of an earlier failed Venus mission. The maximum collision
expectation is determined using the busiest day in 2024 as a representative
proxy of the traffic on the day of re-entry.

<!-- benefits / strengths / weaknesses of the approach --------------------------------------->

The approach presented in this paper builds on previous research.
Our approach utilises H3-tessellation to reduce the spatial complexity.
This cell indexing approach reduces the computation burden significantly.

<!-- future work ---------------------------------------------------------------------------->

Future work will address the derivation of casualty expectation using
seats information for aircraft types in the traffic.
<!-- or can we say something about the next concrete steps to remedy some weaknesses -->

<!-- application of the results -------------------------------------------------------------->

As part of this research and to inform the decision-makers it is planned
to explore the impact of megaconstellations and flight growth forecast and
what would happen in the mid- to long-term in various scenarios of
uncontrolled re-entries and the risk for European airspace.

Of practical interest for the application of this work and future extension
is the scenario of mission failure at launch can typically lead to vast airspace
closure and traffic re-routing: The techniques developed would allow to
quantify risk and impact of the planned measures.

::: {#fig-orbital-elements layout="[40,-4,56]" fig-env="figure*" fig-pos="h" height=80%}
```{r ellipse-parameters}
#| fig-cap: "Keplerian elements: semi-major axis, $a$, the distance between the foci, $2c$, the eccentricity, $e = (2c)/(2a)$. (the semi-parameter (a.k.a. _semilatus rectum_), $p$, describes the size of the conic section by defining the width at the primary focus, and is defined as $p = b^2 / a = a (1 - e^2)$)"

#| echo: false
knitr::include_graphics("media/figures/classical_orbital_elements_2.png")
```

```{r keplerian-elements}
#| fig-cap: "Keplerian elements: inclination, $i$, the right ascension of the ascending node, $\\Omega$ , and the argument of periapsis, $\\omega$.(Earth rotates around $\\hat{K}$)"
#| echo: false

knitr::include_graphics("media/figures/classical_orbital_elements_4.png")
```

Orbital elements, Figures 5-2 and 5-9 from [@sellers2003understanding].
:::

# Appendix: Orbital Elements {#sec-orbitals .appendix}

We refer to @fig-orbital-elements to describe the orbital parameters of
interest for the current work.

The orbital elements describing the trajectory of orbiting object are
illustrated in association with an inertial reference frame
$\hat{I}$-$\hat{J}$-$\hat{K}$. 
The $I$-axis, $\hat{I}$, points toward the vernal equinox and 
the $J$-axis, $\hat{J}$, is perpendicular to the $I$-axis in the equatorial plane
in the sense of Earth's rotation, with the $K$-axis pointing up.
The semi-major axis, $a$, the eccentricity, $e$, the inclination, $i$,
the right ascension of the ascending node, $\Omega$, and the argument of perigee, $\omega$, constitute, together with the mean anomaly at the initial position,
$M_0$, a set of six constants of motion that are preserved in the
absence of perturbation. True anomaly, $\nu$, specifies the location of
an orbiting object within the orbit. It is the angle between perigee and
the orbiting object's position vector measured in the direction of the
orbiting object's motion.

The angular position, $\theta$, of the re-entry debris taken from the ascending
node[^ascending-node] in the orbit plane is $\theta = \omega + \nu$.
It is related to the latitude, $\varphi$, via the Cartesian coordinate $k$,
$k = R \sin(\varphi) = R \sin(\theta)\sin(i)$, or

$$
\sin(\varphi) = \sin(\theta)\sin(i)
$$ {#eq-spheric-coords}

In the derivation of debris orbital positions, we use the value of the mass of Earth, $M_{\oplus} = \qty{5.9722e24}{kg}$ but we neglect that of the debris
because many orders of magnitudes smaller (and not always known).

Of all the orbital elements, for the needs of this paper to numerically compute
the weighting function, , @fig-weighting-inclination-i, we only need
the inclination $i$. 
For debris with inclination $i$ we simulate an orbit of $N$, i.e. $10000$,
positions ($\nu$) and calculate the corresponding Cartesian positions and
the equivalent longitudes and latitudes.
The latitudes are then binned on half degree intervals to compute
$\sigma(\varphi, i)$, as per (@eq-impact-probability-density).

[^ascending-node]: For an earth satellite, the ascending node is the point in its orbit where a satellite crosses the equatorial plane going from south to north.

# References {.unnumbered}

::: {#refs}
:::